---
title: "ACS_data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading the essential libraries for the replication. 

```{r cars}
library(tidyverse)
library(tidycensus)
library(tigris)
options(tigris_use_cache = TRUE)
#install.packages("sf")
library(sf)
library(dplyr)
library(lubridate)
```

User may sign up in the given link in the readme.md file and use the following command to load the ACS data. 
census_api_key("<API_KEY>", install = TRUE)

```{R}
readRenviron("~/.Renviron")
```

Getting the Zip Codes for New York City region. 

```{r}
zip_codes <- read_csv("https://raw.githubusercontent.com/erikgregorywebb/nyc-housing/master/Data/nyc-zip-codes.csv")
  colnames(zip_codes)[which(names(zip_codes)== "ZipCode")] <- "GEOID"

```

Getting the APRIL 01 test and positivity data 

```{r}
april_01 <- read_csv("https://raw.githubusercontent.com/nychealth/coronavirus-data/097cbd70aa00eb635b17b177bc4546b2fce21895/tests-by-zcta.csv")
colnames(april_01)[which(names(april_01)== "MODZCTA")] <- "GEOID"

colnames(april_01)[which(names(april_01)== "Total")] <- "total_tests"

 april_01 <- april_01%>%
   mutate(positivity_prop = Positive/total_tests)

```

FIGURE 1 

Getting the uninsured population aged 18-64

```{r pressure, echo=FALSE}

options(tigris_use_cache = TRUE)


pop_uninsured <- get_acs(geography = "zcta",
                 variables = c(total_18to34_uninsured = 'B27010_018',
                               total_35to64_uninsured = 'B27010_034',
                               pop_uninsured_18to34 ='B27010_033',
                               pop_uninsured_35to64 ='B27010_050'
                              ),
                 state = 'NY',
                 year = 2016,
                 geometry = TRUE)
```

Proportion of 18-64 year olds who are uninsured

```{R}
pop_uninsured<- pop_uninsured %>% merge(zip_codes, by = "GEOID")%>%
  pivot_wider(names_from = variable, values_from = c(estimate, moe))

 pop_uninsured <- pop_uninsured %>% 
   mutate(uninsured_age_bounds = (estimate_pop_uninsured_35to64 + estimate_pop_uninsured_18to34))%>%
   mutate(uninsured_total = (estimate_total_18to34_uninsured + estimate_total_35to64_uninsured))%>%
   mutate(proportion_uninsured = (uninsured_age_bounds/uninsured_total))%>%
   summarize(GEOID, Borough, geometry, uninsured_age_bounds, uninsured_total, proportion_uninsured)

``` 
 
 Replication of the map for Proportion of 18-64 years olds who are uninsured
 
```{R} 
 ggplot(st_as_sf(pop_uninsured), 
        aes(fill = proportion_uninsured))+
   geom_sf()+
   scale_fill_distiller(type = "seq",
                        palette = 8,
                        direction = 1)
``` 
 
 Median calculation for above proportion 
 
```{R} 
 median(pop_uninsured$proportion_uninsured, na.rm = TRUE)
 
```
 
 Forward selection prediction model for above proportion of % COVID positivity by Zip code: April 01, 2020
 
```{r}
pop_uninsured<- pop_uninsured %>% merge(april_01, by = "GEOID")

lm.fit <- lm(positivity_prop ~ proportion_uninsured, data = pop_uninsured)
lm.fit <- lm(positivity_prop ~ proportion_uninsured)

summary(lm.fit)

```

Getting the Median Income (in million, 2016$) 

```{r}

med_income <- get_acs(geography = "zcta",
                 variables = c(
                               med_income = 'B19013_001'
                               ),
                 state = 'NY',
                 year = 2016,
                 geometry = TRUE)
``` 
 
 Merging the Median Income data with the zip code of NYC
```{R} 

med_income<- med_income %>% 
  merge(zip_codes, by = "GEOID")

``` 
 
 Replication of the map for Median income(in millions, $2016)
```{R} 
ggplot(st_as_sf(med_income), 
        aes(fill = estimate))+
   geom_sf()+
   scale_fill_distiller(type = "seq",
                        palette = 9,
                        direction = 1)
``` 
 
 Median calculation of the Median Income
```{R} 

mean(med_income$estimate, na.rm = TRUE)

```

 Forward selection prediction model for above proportion of % COVID positivity by Zip code: April 01, 2020

```{r}
med_income<- med_income %>% 
  merge(april_01, by = "GEOID")
View(med_income)

lm.fit <- lm(positivity_prop ~ estimate, data = med_income)
lm.fit <- lm(positivity_prop ~ estimate)


summary(lm.fit)
```



Getting the population of people self-identifying as White. 

```{r}
pop_white <- get_acs(geography = "zcta",
                 variables = c(total_race_pop = 'B02001_001',
                               prop_white = 'B02001_002'),
                 state = 'NY',
                 year = 2016,
                 geometry = TRUE)
```

Merging the white population with the NYC ZipCode
```{R}
pop_white <- pop_white %>%
  merge(zip_codes, by = "GEOID")%>%
  pivot_wider(names_from = variable, values_from = c(estimate, moe))

```

Proportion of self-identifying as White
```{R}
pop_white <- pop_white %>%
  mutate(white_pop_prop = estimate_prop_white/estimate_total_race_pop )%>%
  summarize(GEOID, Borough, geometry, estimate_prop_white, estimate_total_race_pop, white_pop_prop)

```

Replicating the map for proportion self-identifying as White
```{R}
ggplot(st_as_sf(pop_white), 
        aes(fill = white_pop_prop))+
   geom_sf()+
   scale_fill_distiller(type = "seq",
                        palette = 10,
                        direction = 1)
```

Median of Proportion of self-identifying as White

```{R}
median(pop_white$white_pop_prop, na.rm = TRUE)

```



Forward selection prediction model for above proportion of % COVID positivity by ZipCode: April 01, 2020

```{r}
pop_white<- pop_white %>% 
  merge(april_01, by = "GEOID")

lm.fit <- lm(positivity_prop.x ~ white_pop_prop, data = pop_white)
lm.fit <- lm(positivity_prop.x ~ white_pop_prop)

lm.fit

summary(lm.fit)
```


Getting the population of Households of four or more 

```{r}
pop_household <- get_acs(geography = "zcta",
                 variables = c(pop_household_size = 'B11016_001',
                               pop_hsh_above_three_4 = 'B11016_005',
                               pop_hsh_above_three_5 ='B11016_006',
                               pop_hsh_above_three_6 ='B11016_007',
                               pop_hsh_above_three_7 ='B11016_008'),
                 state = 'NY',
                 year = 2016,
                 geometry = TRUE)
```

Sorting the population on the basis of NYC ZipCode
```{r}
pop_household <- pop_household %>%
   merge(zip_codes, by = "GEOID")%>%
  pivot_wider(names_from = variable, values_from = c(estimate, moe))
```

Proportion in housholds of 4 or more. 
```{r}
pop_household <- pop_household %>%
  mutate(total_household_above_three = estimate_pop_hsh_above_three_4+estimate_pop_hsh_above_three_5+estimate_pop_hsh_above_three_6+estimate_pop_hsh_above_three_7)%>%
  mutate(proportion_household = total_household_above_three/estimate_pop_household_size)%>%
  summarize(GEOID, Borough, geometry, total_household_above_three, estimate_pop_household_size, proportion_household)

```

Median of Proportion in housholds of 4 or more.
```{r}

median(pop_household$proportion_household, na.rm = TRUE)
```

Replicating the map for Proportion in housholds of 4 or more.

```{r}
ggplot(st_as_sf(pop_household), 
        aes(fill = proportion_household))+
   geom_sf()+
   scale_fill_distiller(type = "seq",
                        palette = 11,
                        direction = 1)
```

Forward selection prediction model for above proportion of % COVID positivity by ZipCode: April 01, 2020

```{r}
pop_household <- pop_household %>% 
  merge(april_01, by = "GEOID")

lm.fit <- lm(positivity_prop.x ~ proportion_household, data = pop_household)
lm.fit <- lm(positivity_prop.x ~ proportion_household)

summary(lm.fit)
```

Getting the population used bus as public transportation

```{r}
public_transportation <- get_acs(geography = "zcta",
                 variables = c(pop_transportation = 'B08301_001',
                               pop_trans_bus = 'B08301_011'),
                 state = 'NY',
                 year = 2016,
                 geometry = TRUE)
```

Sorting the population by the NYC ZipCodes. 
```{r}
public_transportation <- public_transportation %>%
  merge(zip_codes, by = "GEOID")%>%
  pivot_wider(names_from = variable, values_from = c(estimate, moe))
```

Proportion of pppulation that commutes by bus 
```{R}
public_transportation <- public_transportation %>%
  mutate(proportion_trans = estimate_pop_trans_bus/estimate_pop_transportation)
```

Replicating the map for Proportion of population that commutes by bus 
```{r}
ggplot(st_as_sf(public_transportation), 
        aes(fill = proportion_trans))+
   geom_sf()+
   scale_fill_distiller(type = "seq",
                        palette = 13,
                        direction = 1)
```

Median of Proportion of population that commutes by bus 

```{r}
median(public_transportation$proportion_trans, na.rm=TRUE)


```

Forward selection prediction model for above proportion of % COVID positivity by ZipCode: April 01, 2020

```{r}
public_transportation <- public_transportation %>% 
  merge(april_01, by = "GEOID")

lm.fit <- lm(positivity_prop.x ~ proportion_trans, data = public_transportation)
lm.fit <- lm(positivity_prop.x ~ proportion_trans)

summary(lm.fit)
```

Getting the population 65+ years of age

```{r}

pop_above65 <- get_acs(geography = "zcta",
                 variables = c(pop_above_65 = 'B01001_001',
                               Mpop_popln_65_66 = 'B01001_020',
                               Mpop_popln_67_69 ='B01001_021',
                               Mpop_popln_70_74 ='B01001_022',
                               Mpop_popln_75_79 ='B01001_023',
                               Mpop_popln_80_84 ='B01001_024',
                               Mpop_popln_above_85 ='B01001_025',
                               Fpop_popln_65_66 ='B01001_044',
                               Fpop_popln_67_69 ='B01001_045',
                               Fpop_popln_70_74 ='B01001_046',
                               Fpop_popln_75_79 = 'B01001_047',
                               Fpop_popln_80_84 ='B01001_048',
                               Fpop_popln_above_85 ='B01001_049'),
                 state = 'NY',
                 year = 2016,
                 geometry = TRUE)
```

Sorting the population by NYC Zipcodes. 
```{r}
pop_above65 <- pop_above65 %>%
  merge(zip_codes, by = "GEOID")%>%
pivot_wider(names_from = variable, values_from = c(estimate, moe))

```

Proportion of population 65+ years of age
```{r}

pop_above65 <- pop_above65 %>%
  
  mutate(total_popn = estimate_Fpop_popln_65_66+estimate_Fpop_popln_67_69+estimate_Fpop_popln_70_74+estimate_Fpop_popln_75_79+estimate_Fpop_popln_80_84+estimate_Fpop_popln_above_85+estimate_Mpop_popln_65_66+estimate_Mpop_popln_70_74+estimate_Mpop_popln_67_69 + estimate_Mpop_popln_75_79 + estimate_Mpop_popln_80_84 + estimate_Mpop_popln_above_85)%>%
 
  mutate(proportion_above65 = total_popn/estimate_pop_above_65)%>%
  
  summarize(GEOID, Borough, geometry, total_popn , estimate_pop_above_65, proportion_above65)

```

Median of Proportion of population 65+ years of age
```{r}

median(pop_above65$proportion_above65, na.rm=TRUE)

```

Replicating the map for Proportion of population 65+ years of age
```{r}
ggplot(st_as_sf(pop_above65), 
        aes(fill = proportion_above65))+
   geom_sf()+
   scale_fill_distiller(type = "seq",
                        palette = 14,
                        direction = 1)

```

*Replication of Table 1 for April 01, 2020*

```{r}
lm.fit <- lm(positivity_prop ~ proportion_household + proportion_uninsured + white_pop_prop + estimate)
summary(lm.fit)

```



FIGURE 2 

Loading the SafeGraph data
```{r}
load('/data/safegraph/safegraph.Rdata')

safegraph<- safegraph%>%
  rename(GEOID = postal_code)%>%
  merge(zip_codes, by = "GEOID")

```


Getting the data for February from SafeGraph
```{r}
feb_data <- safegraph%>%
  filter(month(date)==2) %>%
  arrange(GEOID, date)

```


Median of February Data 

```{r}
med_feb <- feb_data %>%
  group_by(GEOID)%>%
  summarize(median = median(avg_visits_per_day))

```

Getting the data for March data 

```{r}
mar_data <- safegraph%>%
  filter(month(date)==3) %>%
  arrange(GEOID, date)
```


Getting the data for April data 

```{r}
apr_data <- safegraph%>%
  filter(month(date)==4)%>%
  arrange(GEOID, date)

```

Mobility of March and April data 

```{r}
mobility <- mar_data %>%
  rbind( apr_data)%>%
  merge(med_feb, by = "GEOID")%>%
  mutate(prop_of_moblt = (avg_visits_per_day- median)/median)

mobility <- mobility%>%
  filter(abs(prop_of_moblt) <= 1)
View(mobility)
mobility_quantile <- mobility%>%
  group_by(date)%>%
  summarise(
 q3 = quantile(prop_of_moblt, probs = 0.75, na.rm = TRUE),
  q1 = quantile(prop_of_moblt, probs = 0.25, na.rm = TRUE),
  median = quantile(prop_of_moblt, probs = 0.5, na.rm = TRUE))

```


Replicating the figure 2 for change inmobility by date. 

```{r}

library(lubridate)
 
ggplot()+
  geom_violin(data = mobility, aes(x = prop_of_moblt,
                            y = as.factor(date)),
                  color = "Orange")+
  geom_pointrange(data = mobility_quantile, aes(xmin = q1, 
                                         xmax= q3, 
                                         x = median,
                                         y = as.factor(date)), 
                  color= "Red")
  
```


Replication of Table 2 - April 01, 2020 

No mobility

```{r}
mobility <- mobility %>% 
  merge(april_01, by = "GEOID")


lm_no_mobility <- lm(positivity_prop ~ proportion_household + proportion_uninsured + white_pop_prop + estimate, data = mobility)

summary(lm_no_mobility)
```

With mobility 
```{r}

lm_mobility <- lm(positivity_prop ~ proportion_household + proportion_uninsured + white_pop_prop + estimate+ prop_of_moblt, data = mobility)


summary(lm_mobility)
```





















